Information-theoretic measures for non-linear causality detection: \\ application to social media sentiment and cryptocurrency prices

### 2020-07-25 paper/submission.tex
Removed blue highlighting for final .tex version v2

  In modern portfolio theory, investors commonly calculate correlations between asset types to construct portfolios aiming to maximise their return at a given level of risk \cite{markowitz1952portfolio}. In the search for excess returns, quantitative approaches are often exploited to detect predictive signals across time series. This may be considered as a field of price discovery, which represents a large and fundamental area of financial theory. In the ideal case {with complete predictive information, \color{blue}from} knowing the movement of one price, the movement of a second can be inferred exactly. In reality information is incomplete and held asymmetrically, and the ability to reliably obtain this and contribute to price formation is generally rewarded by the market. In this paper we explore the effectiveness of two promising techniques for detecting information transfer between time series, applying these to identify causal signals between alternative data and cryptocurrency price returns.

@@

  where $X_t$ is the value of $X$ at time $t$, and $r$ is a parameter which in fact defines the dynamical state of the system. {\color{black}Following Hahs \& Pethel \cite{hahs2011distinguishing}, we take $r=4$ so the function evolves chaotically.} We then introduce a second map, which is dependent on the first, taking the form:

@@

 where $\epsilon \in [0,1]$ represents the coupling strength, describing the extent to which $Y_{t+1}$ depends on $f(f(X_t))$. It should be noted that the logistic map, in contrast to geometric Brownian motion, is a deterministic, albeit chaotic system, and that therefore $f(f(X_t))$ is equivalent to $X_{t+2}$. {\color{black} The extent of this anticipatory effect is driven by the selection of the $\epsilon$ parameter. We follow Hahs \& Pethel in selecting $\epsilon=0.4$. Indeed, as $\alpha$ increases, with large $\epsilon$, the direction of information transfer is less clear, as $Y_t$ contains more information about the future values of $X$. }

>>>>

  In modern portfolio theory, investors commonly calculate correlations between asset types to construct portfolios aiming to maximise their return at a given level of risk \cite{markowitz1952portfolio}. In the search for excess returns, quantitative approaches are often exploited to detect predictive signals across time series. This may be considered as a field of price discovery, which represents a large and fundamental area of financial theory. In the ideal case {with complete predictive information, from knowing the movement of one price, the movement of a second can be inferred exactly. In reality information is incomplete and held asymmetrically, and the ability to reliably obtain this and contribute to price formation is generally rewarded by the market. In this paper we explore the effectiveness of two promising techniques for detecting information transfer between time series, applying these to identify causal signals between alternative data and cryptocurrency price returns.

@@

  where $X_t$ is the value of $X$ at time $t$, and $r$ is a parameter which in fact defines the dynamical state of the system. Following Hahs \& Pethel \cite{hahs2011distinguishing}, we take $r=4$ so the function evolves chaotically. We then introduce a second map, which is dependent on the first, taking the form:

@@

 where $\epsilon \in [0,1]$ represents the coupling strength, describing the extent to which $Y_{t+1}$ depends on $f(f(X_t))$. It should be noted that the logistic map, in contrast to geometric Brownian motion, is a deterministic, albeit chaotic system, and that therefore $f(f(X_t))$ is equivalent to $X_{t+2}$. The extent of this anticipatory effect is driven by the selection of the $\epsilon$ parameter. We follow Hahs \& Pethel in selecting $\epsilon=0.4$. Indeed, as $\alpha$ increases, with large $\epsilon$, the direction of information transfer is less clear, as $Y_t$ contains more information about the future values of $X$.
